% -*- TeX-engine: xetex; eval: (auto-fill-mode 0); eval: (visual-line-mode 1); -*-
% Compile with XeLaTeX

%%%%%%%%%%%%%%%%%%%%%%%
% Option 1: Slides: (comment for handouts)   %
%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[slidestop,compress,mathserif,12pt,t,professionalfonts,xcolor=table]{beamer}

% solution stuff
\newcommand{\solnMult}[1]{
\only<1>{#1}
\only<2->{\red{\textbf{#1}}}
}
\newcommand{\soln}[1]{\textit{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Option 2: Handouts, without solutions (post before class)    %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \documentclass[11pt,containsverbatim,handout,xcolor=xelatex,dvipsnames,table]{beamer}
%
% % handout layout
% \usepackage{pgfpages}
% \pgfpagesuselayout{4 on 1}[letterpaper,landscape,border shrink=5mm]
%
%% % solution stuff
% \newcommand{\solnMult}[1]{#1}
% \newcommand{\soln}[1]{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Option 3: Handouts, with solutions (may post after class if need be)    %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \documentclass[11pt,containsverbatim,handout,xcolor=xelatex,dvipsnames,table]{beamer}

% % handout layout
% \usepackage{pgfpages}
% \pgfpagesuselayout{4 on 1}[letterpaper,landscape,border shrink=5mm]

% % solution stuff
% \newcommand{\solnMult}[1]{\red{\textbf{#1}}}
% \newcommand{\soln}[1]{\textit{#1}}

%%%%%%%%%%
% Load style file, defaults  %
%%%%%%%%%%

\input{../../lec_style.tex}
\input{../../definitions_default.tex}
% ALT ALT
% \input{../../definitions_custom.tex}

%%%%%%%%%%%
% Cover slide info    %
%%%%%%%%%%%

\title{Unit 5: Inference for categorical data}
\subtitle{1. Inference for a single proportion}
\author{\CourseName}
\date{}
\institute{\InstituteName}


%%%%%%%%%%%%%%%%%%%%%%%%%
% Begin document and set Helvetica Neue font   %
%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\fontspec[Ligatures=TeX]{Helvetica Neue Light}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Title Page

\begin{frame}[plain]

\titlepage

\vfill

{\scriptsize \webLink{\PersonalSite}{Dr. \LastName{}} \hfill Slides posted at  \webURL{\CourseSite}}

\addtocounter{framenumber}{-1} 

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Housekeeping}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Announcements}

\begin{itemize}

\item PS 5 due April 1 (after the midterm) but I strongly recommend that you work on it and ask questions before.

\item PA 5 is due Mar 29 (so you can review it before the midterm). It will be available on Mar 23 but will contain material on chi-square testing which we will cover on Mar 28, so you can decide when you want to take it.

\item Sunday, Mar 27, 2-3pm at Old Chem 116.

\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Main ideas}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{The CLT also describes the distribution of $\hat{p}$}
\label{mi2}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Distribution of $\hat{p}$}

\hl{Central limit theorem for proportions:} Sample proportions will be nearly normally distributed with 
mean equal to the population mean, $p$, and standard error equal to $\sqrt{\frac{p~(1-p)}{n}}$.
\[ \hat{p} \sim N \pr{ mean = p, SE = \sqrt{\frac{p~(1-p)}{n}} } \]

Conditions:
\begin{itemize}
\item Independence: Random sample/assignment + 10\% rule
\item At least 10 successes and failures
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{}

\clicker{Suppose $p = 0.05$. What shape does the distribution of $\hat{p}$ have in random samples of $n = 100$.}

\begin{enumerate}[(a)]
\item unimodal and symmetric (nearly normal)
\item bimodal and symmetric
\item \solnMult{right skewed}
\item left skewed
\end{enumerate}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{}

\clicker{Suppose $p = 0.5$. What shape does the distribution of $\hat{p}$ have in random samples of $n = 100$.}

\begin{enumerate}[(a)]
\item \solnMult{unimodal and symmetric (nearly normal)}
\item bimodal and symmetric
\item right skewed
\item left skewed
\end{enumerate}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{CI vs. HT determines observed vs. expected counts / proportions}
\label{mi3}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{CI vs. HT determines observed vs. expected counts / proportions}

Remember, when doing a HT always assume $H_0$ is true!
\pause

\begin{itemize}

\item \textbf{S-F:} Number of successes and failures for checking the success-failure condition for the 
nearly normal distribution of $\hat{p}$:
\pause
\begin{itemize}
\item CI: use observed proportion $\rightarrow$ $n\hat{p} \ge 10$ and $n(1 - \hat{p}) \ge 10$
\pause
\item HT: use null value of the proportion $\rightarrow$ $np_0 \ge 10$ and $n(1 - p_0) \ge 10$
\end{itemize}

\pause

\item \textbf{SE:} Proportion of success for calculating the standard error of $\hat{p}$: 
\[ SE = \sqrt{\frac{p(1-p)}{n}} \]
\pause
\vspace{-0.5cm}
\begin{itemize}
\item CI: use observed proportion $\rightarrow$ $SE = \sqrt{\frac{\hat{p}(1-\hat{p})}{n}}$
\pause
\item HT: use null value of the proportion $\rightarrow$ $SE = \sqrt{\frac{p_0(1-p_0)}{n}}$
\end{itemize}

\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Only use CLT based methods if the sample size is large enough for a nearly normal sampling distribution}
\label{mi4}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Simulation vs. theoretical inference}

\begin{itemize}

\item If the S-F condition is met, can do theoretical inference: Z test, Z interval

\item If the S-F condition is not met, must use simulation based methods: randomization test, bootstrap interval

\end{itemize}

\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Applications}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Single population proportion, large sample}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\vfill

\app{App Ex 5.1}{See course website for details.}

\vfill

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Single population proportion, small sample}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\clicker{Are you vegetarian or vegan?}

\begin{enumerate}[(a)]
\item Yes, I am vegetarian or vegan
\item No, I am neither vegetarian nor vegan
\end{enumerate}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}

\clicker{A variety of studies suggest that 8\% of college students are vegetarian or vegan. Assuming that this 
class is a representative sample of Duke students, which of the following are the correct set of hypotheses 
for testing if the proportion of Duke students who are vegetarian is different than the proportion of 
vegetarian college students at large.}

\begin{enumerate}[(a)]
\item \solnMult{$H_0: p = 0.08$; $H_A: p \ne 0.08$}
\item $H_0: p = 0.08$; $H_A: p < 0.08$
\item $H_0: \hat{p} = 0.08$; $H_A: \hat{p} \ne 0.08$
\item $H_0: \hat{p}_{Duke} = \hat{p}_{all~college}$; $H_A: \hat{p}_{Duke} \ne \hat{p}_{all~college}$
\item $H_0: p_{Duke} = p_{all~college}$; $H_A: p_{Duke} \ne p_{all~college}$
\end{enumerate}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Simulate by hand}

\disc{Describe a simulation scheme for this hypothesis test.}

\pause

\begin{itemize}
\item 100 chips in a bag: 8 green (veg), 92 white (non veg).
\pause
\item Sample randomly $n$ times from the bag, with replacement ($n$ = observed sample size)
\pause
\item Calculate $\hat{p}$, the proportion of greens (successes) in the random sample of size $n$, 
record this value.
\pause
\item Repeat many times.
\pause
\item Calculate the proportion of simulations where $\hat{p}$ is at least as different from 0.08 
as the observed sample proportion.
\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]
\frametitle{HT in R}

{\scriptsize
\begin{Verbatim}[frame=single, formatcom=\color{blue}]
n_veg = [fill in based on class data]
n_nonveg = [fill in based on class data]

sta101 = data.frame(veg = c(rep("yes", n_veg), rep("no", n_nonveg)))

inference(y = veg, data = sta101, success = "yes", 
          statistic = "proportion", type = "ht", 
          null = 0.08, alternative = "twosided", 
          method = "simulation")
\end{Verbatim}
}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Bootstrap interval for a single proportion}

\vfill

\disc{How would the simulation scheme change for a bootstrap interval for the proportion of Duke 
students who are vegetarians?}

\vfill

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}[fragile]
\frametitle{CI in R}

{\scriptsize
\begin{Verbatim}[frame=single, formatcom=\color{blue}]
inference(y = veg, data = sta101, success = "yes", 
          statistic = "proportion", type = "ci", 
          method = "simulation", boot_method = "se")
\end{Verbatim}
}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Recap}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Recap on CLT based methods}

\begin{itemize}

\item Calculating the necessary sample size for a CI with a given margin of error:
\begin{itemize}
\item If there is a previous study, use $\hat{p}$ from that study
\item If not, use $\hat{p} = 0.5$:
\begin{itemize}
\item if you don't know any better, 50-50 is a good guess
\item $\hat{p} = 0.5$ gives the most conservative estimate -- highest possible sample size
\end{itemize}
\end{itemize}

\pause

\item HT vs. CI for a proportion
\begin{itemize}

\item Success-failure condition:
\begin{itemize}
\item CI: At least 10 \red{observed} successes and failures
\item HT: At least 10 \red{expected} successes and failures, calculated using the null value
\end{itemize}

\item Standard error:
\begin{itemize}
\item CI: calculate using observed sample proportion: $SE = \sqrt{\frac{p(1-p)}{n}}$
\item HT: calculate using the null value: $SE = \sqrt{\frac{p_0(1-p_0)}{n}}$
\end{itemize}

\end{itemize}


\end{itemize}

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Recap on simulation methods}

If the S-F condition is not met

\begin{itemize}

\item HT: Randomization test -- simulate under the assumption that $H_0$ is true, then find the p-value as 
proportion of simulations where the simulated $\hat{p}$ is at least as extreme as the one observed.

\item CI: Bootstrap interval -- resample with replacement from the original sample, and construct interval 
using percentile or standard error method.

\end{itemize}


\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Summary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
\frametitle{Summary of main ideas}

\vfill

\begin{enumerate}

\item \nameref{mi2}

\item \nameref{mi3}

\item \nameref{mi4}

\end{enumerate}

\vfill

\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}