---
title: 'Foundations for statistical inference - Confidence intervals'
output:
  html_document:
    css: ../lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
library(dplyr)
library(ggplot2)
library(mosaic)
set.seed(9438024)

normTail <- function (m = 0, s = 1, L = NULL, U = NULL, M = NULL, df = 1000, 
    curveColor = 1, border = 1, col = "#CCCCCC", xlim = NULL, 
    ylim = NULL, xlab = "", ylab = "", digits = 2, axes = 1, 
    detail = 999, xLab = c("number", "symbol"), cex.axis = 1, 
    xAxisIncr = 1, ...) 
{
    if (is.null(xlim)[1]) {
        xlim <- m + c(-1, 1) * 3.5 * s
    }
    temp <- diff(range(xlim))
    x <- seq(xlim[1] - temp/4, xlim[2] + temp/4, length.out = detail)
    y <- dt((x - m)/s, df)/s
    if (is.null(ylim)[1]) {
        ylim <- range(c(0, y))
    }
    #cat(x)
    #cat(y)
    plot(x, y, type = "l", xlim = xlim, ylim = ylim, xlab = xlab, 
        ylab = ylab, axes = FALSE, col = curveColor, ...)
    if (!is.null(L[1])) {
        these <- (x <= L)
        X <- c(x[these][1], x[these], rev(x[these])[1])
        Y <- c(0, y[these], 0)
        polygon(X, Y, border = border, col = col)
    }
    if (!is.null(U[1])) {
        these <- (x >= U)
        X <- c(x[these][1], x[these], rev(x[these])[1])
        Y <- c(0, y[these], 0)
        polygon(X, Y, border = border, col = col)
    }
    if (all(!is.null(M[1:2]))) {
        these <- (x >= M[1] & x <= M[2])
        X <- c(x[these][1], x[these], rev(x[these])[1])
        Y <- c(0, y[these], 0)
        polygon(X, Y, border = border, col = col)
    }
    if (axes == 1 || axes > 2) {
        if (xLab[1] == "symbol") {
            xAt <- m + (-3:3) * s
            xLab <- expression(mu - 3 * sigma, mu - 2 * sigma, 
                mu - sigma, mu, mu + sigma, mu + 2 * sigma, mu + 
                  3 * sigma)
        }
        else if (xLab[1] != "number") {
            stop("Argument \"xLab\" not recognized.\n")
        }
        else {
            temp <- seq(xAxisIncr, max(abs(xlim - m))/s, xAxisIncr) * 
                s
            xAt <- m + c(-temp, 0, temp)
            xLab <- round(xAt, digits = digits)
        }
    }
    if (axes > 2) {
        axis(1, at = xAt, labels = xLab, cex.axis = cex.axis)
        buildAxis(2, c(y, 0), n = 3, nMax = 3, cex.axis = cex.axis)
    }
    else if (axes > 1) {
        buildAxis(2, c(y, 0), n = 3, nMax = 3, cex.axis = cex.axis)
    }
    else if (axes > 0) {
        axis(1, at = xAt, labels = xLab, cex.axis = cex.axis)
    }
    abline(h = 0)
}
```

<div id="template">
**Your reproducible lab report:** Before you get started, download the R Markdown 
template for this lab. Remember all of your code and answers go in this document:
```{r load-template, eval = FALSE}
download("http://stat.duke.edu/courses/Fall15/sta101.002/rmd/lab5.Rmd", 
         destfile = "lab5.Rmd")
```
</div>

## Sampling from Ames, Iowa

If you have access to data on an entire population, say the size of every 
house in Ames, Iowa, it's straight forward to answer questions like, "How big 
is the typical house in Ames?" and "How much variation is there in sizes of 
houses?". If you have access to only a sample of the population, as is often 
the case, the task becomes more complicated. What is your best guess for the 
typical size if you only know the sizes of several dozen houses? This sort of 
situation requires that you use your sample to make inference on what your 
population looks like.

<div id="boxedtext">
**Setting a seed:** We will take some random samples and calculate confidence based
on these samples in this lab, which means you should set a seed on top of your lab. If 
this concept is new to you, review the previous lab and ask your TA.
</div>

## The data

In the previous lab, ``Sampling Distributions'', we looked at the population data
of houses from Ames, Iowa. Let's start by loading that data set.

```{r load-data}
load(url("https://stat.duke.edu/~mc301/data/ames.RData"))
```

In this lab we'll start with a simple random sample of size 60 from the 
population. Specifically, this is a simple random sample of size 60. Note that 
the data set has information on many housing variables, but for the first 
portion of the lab we'll focus on the size of the house, represented by the 
variable `area`.

```{r sample}
n <- 60
samp <- sample_n(ames, n)
```

1.  Describe the distribution of homes in your sample. What would you say is the
    "typical" size within your sample? Also state precisely what you interpreted 
    "typical" to mean.

1.  Would you expect another team's distribution to be identical to yours? 
    Would you expect it to be similar? Why or why not?

## Confidence intervals

Return for a moment to the question that first motivated this lab: based on 
this sample, what can we infer about the population? Based only on this single 
sample, the best estimate of the average living area of houses sold in Ames 
would be the sample mean, usually denoted as $\bar{x}$ (here we're calling it 
`x_bar`). That serves as a good **point estimate** but it would be useful 
to also communicate how uncertain we are of that estimate. This uncertainty
can be quantified using a **confidence interval**.

A confidence interval for a population mean is of the following form
\[ \bar{x} + z^\star \frac{s}{\sqrt{n}} \]

You should by now be comfortable with calculating the mean and standard deviation of 
a sample in R. And we know that the sample size is 60. So the only remaining building
block is finding the appropriate critical value for a given confidence level. We can
use the `qnorm` function for this task, which will give the critical value associated
with a given percentile under the normal distribution. Remember that confidence levels
and percentiles are not equivalent. For example, a 95% confidence level refers to the
middle 95% of the distribution, and the critical value associated with this area will
correspond to the 97.5th percentile.

```{r eval=TRUE, echo=FALSE, fig.height=3, fig.width=5, fig.align='center'}
par(mar = c(3,0,0,0))
normTail(M = c(-1.96, 1.96), col = "#569BBD", axes = FALSE)
axis(1, at = c(-3, -1.96, 0, 1.96, 3), labels = c(NA, -1.96, 0, 1.96, NA), 
     cex.axis = 1.5)
text(x = 0, y = 0.15, "95%", cex = 1.5, col = "white")
text(x = -2.8, y = 0.05, "2.5%", cex = 1.5)
text(x = 2.8, y = 0.05, "2.5%", cex = 1.5)
graphics::arrows(-3.7, 0.25, 1.96, 0.25, col = "red", lwd = 3, code = 3)
text(x = -1.4, y = 0.275, "97.5%", cex = 1.5, col = "red")
```

We can find the critical value for a 95% confidence interal using
```{r z_star_95}
z_star_95 <- qnorm(0.975)
z_star_95
```
which is roughly equal to the value critical value 1.96 that you're likely
familiar with by now.

Let's finally calculate the confidence interval:
```{r ci}
samp %>%
  summarise(lower = mean(area) - z_star_95 * (sd(area) / sqrt(n)),
            upper = mean(area) + z_star_95 * (sd(area) / sqrt(n)))
```

To recap: even though we don't know what the full population looks like, we're 95% 
confident that the true average size of houses in Ames lies between the values *lower* 
and *upper*. There are a few conditions that must be met for this interval to be valid.

1.  For the confidence interval to be valid, the sample mean must be normally 
    distributed and have standard error $s / \sqrt{n}$. What conditions must be 
    met for this to be true?

## Confidence levels

1.  What does "95% confidence" mean?

In this case we have the rare luxury of knowing the true population mean since we 
have data on the entire population. Let's calculate this value so that
we can determine if our confidence intervals actually capture it. We'll store it in a
data frame called `params` (short for population parameters), and name it `mu`.

```{r pop-mean}
params <- ames %>%
  summarise(mu = mean(area))
```

1.  Does your confidence interval capture the true average size of houses in 
    Ames? If you are working on this lab in a classroom, does your neighbor's 
    interval capture this value? 

1.  Each team in should have gotten a slightly different confidence interval. What 
    proportion of those intervals would you expect to capture the true population 
    mean? Why?

Using R, we're going to collect many samples to learn more about how sample 
means and confidence intervals vary from one sample to another.

Here is the rough outline:

-   Obtain a random sample.
-   Calculate the sample's mean and standard deviation, and use these to calculate
and store the lower and upper bounds of the confidence intervals.
-   Repeat these steps 50 times.

We can accomplish this using the `do` function. The following lines of code takes 50
random samples of size `n` from population (and remember we defined $n = 60$ earlier),
and computes the upper and lower bounds of the confidence intervals based on these
samples.

```{r calculate-50-cis}
ci <- do(50) * ames %>%
                  sample_n(n) %>%
                  summarise(lower = mean(area) - z_star_95 * (sd(area) / sqrt(n)),
                            upper = mean(area) + z_star_95 * (sd(area) / sqrt(n)))
```

Let's view the first five intervals:

```{r first-five-intervals}
ci %>%
  slice(1:5)
```

Next we'll create a plot similar to Figure 4.8 on page 175 of [OpenIntro Statistics, 3rd
Edition](https://www.openintro.org/os). First step will be to create a new variable in 
the `ci` data frame that indicates whether the interval does or does not capture the 
true population mean. Note that capturing this value would mean the lower bound of the
confidence interval is below the value and upper bound of the confidence interval is
above the value. Remember that we create new variables using the `mutate` function.

```{r capture-mu}
ci <- ci %>%
  mutate(capture_mu = ifelse(lower < params$mu & upper > params$mu, "yes", "no"))
```

The `ifelse` function is new. It takes three arguments: first is a logical statement,
second is the value we want if the logical statement yields a true result, and the
third is the value we want if the logical statement yields a false result.

We now have all the information we need to create the plot, but we need to re-organize
our data a bit for easy plotting. Specifically, we need to organize the data in a new
data frame where each row represents one bound, as opposed to one interval. So this

~~~
     lower    upper capture_mu
1 1355.688 1559.545        yes
2 1413.661 1681.673        yes
3 1372.159 1631.807        yes
...
~~~

should instead look like

~~~
  ci_id ci_bounds capture_mu
1     1  1355.688        yes
2     2  1413.661        yes
3     3  1372.159        yes
4     1  1559.545        yes
5     2  1681.673        yes
6     3  1631.807        yes
...
~~~

We can accomplish this using the following:

```{r create-ci-data-for-plot}
ci_data <- data.frame(ci_id = c(1:50, 1:50),
                      ci_bounds = c(ci$lower, ci$upper),
                      capture_mu = c(ci$capture, ci$capture))
```

And finally we can create the plot using the following:

```{r plot-ci}
qplot(data = ci_data, x = ci_bounds, y = ci_id, 
      group = ci_id, color = capture_mu) +
  geom_point(size = 2) +  # add points at the ends, size = 2
  geom_line() +           # connect with lines
  geom_vline(xintercept = params$mu, color = "darkgray") # draw vertical line
```

1.  What proportion of your confidence intervals include the true population mean? Is 
    this proportion exactly equal to the confidence level? If not, explain why. Make 
    sure to include your plot in your answer.

* * *

<div class="oyo">

## On your own

-   Pick a confidence level of your choosing, provided it is not 95%. What is 
    the appropriate critical value?

-   Calculate 50 confidence intervals at the confidence level you chose in the 
    previous question, and plot all intervals on one plot, and calculate the proportion 
    of intervals that include the true population mean. How does this percentage compare 
    to the confidence level selected for the intervals? Make 
    sure to include your plot in your answer.
</div>

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was written for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel.
</div>